

\newpage
\section{A1.4) Exercise 5.13}

\begin{itemize}
  \item \emph{Solve exercise 5.13 from the course compendium.}
\end{itemize}

\textcolor{red}{In my feedback, I was asked to name more advantages and
disadvantages to using between using a heap vs. stack. The newly added answers
are in red.}

\medskip

A frame contains parameters and (space for) local variables to the given
function, as well as a return address for the function. Most functions will not
have very many parameters and local variables - perhaps 3-6 and 5-15,
respectively - and therefore most frames are relatively small. This plays well
with generational collection, since in most cases the entire list of frames in
play can fit in the first generation (or the first couple of generations), and
can reduce the time taken for frame deallocation. \smallskip

However, if a program exhibits many consecutive function calls, such as in the loop
below:

\begin{minted}[linenos=false]{c}
for (i = 0; i < n; i++) {
  acc = f(acc, input[i]);
}
\end{minted}

then the lower generations of the heap may repeatedly be filled and consequently
collected many times throughout the loop; in the worst case, the lower
generations might be rendered entirely useless for the duration of the loop,
since actual user data that is to stay alive throughout the loop is quickly
being moved to higher generations to make room for frames in these lower
generations. This may to some degree nullify the benefit from generational GC.

\medskip

\paragraph{\textcolor{red}{Other considerations}}~\smallskip

\textcolor{red}{There are of course other things to consider besides the effects
on (generational) garbage collection. For one, one very big disadvantage in
allocating frames in the heap is the problem of fragmentation, which can never
be a problem with stack allocated frames.}

\medskip

\textcolor{red}{On the other hand, allocating frames on the heap has the big
advantage that whereas there is typically a static bound on the size of stack
memory, the heap size is only bounded by the amount of available memory or by
how much the operating system is willing to allocate for the program. This
allows for potentially much deeper recursion, which may or may not be a benefit
in some programs.}

\Sectend
