\newpage
\section{ARIES}

\begin{itemize}
  \item \textit{1. The transaction and dirty page tables after the analysis
    phase.}
\end{itemize}


\begin{minted}{text}
/-----------------------------\
|      Transaction Table      |
|-----------------------------|
| xact id | status | last_LSN |
|---------+--------+----------|
| T1      | U      | 9        |
\-----------------------------/

/--------------------\
|  Dirty Page Table  |
|--------------------|
| page id | rec_LSN  |
|---------|----------|
| P2      | 3        |
| P1      | 4        |
| P5      | 5        |
\--------------------/
\end{minted}

The last log record for \ms{T2} was a commit, so it would have been temporarily
inserted into the transaction table during the analysis phase, but afterwards it
would have been removed and an \ms{end} record written to the log.
\streg

\begin{itemize}
  \item \textit{2. The sets of winner and loser transactions.}
\end{itemize}

The set of winner transaction is simply the singleton set \ms{{T3}}, while the
loser transactions is \ms{{T1, T2}}, since these were still active at the time
of crash.

\streg

\begin{itemize}
  \item \textit{3. The LSNs for the start of the redo phase and the end of the
    undo phase.}
\end{itemize}

The redo phase starts at \ms{LSN = 3}, since this is the smallest \ms{recLSN} in
the dirty page table, and thus the earliest log record which could possibly have
to be redone.
\smallskip

The undo phase also ends at \ms{LSN = 3}, since in undoing transaction \ms{T1}
this is the earliest \ms{LSN} we will encounter.

\streg

\begin{itemize}
  \item \textit{4. The set of log records that may cause pages to be rewritten
    during the redo phase.}
\end{itemize}

The set in question is \ms{{3, 4, 5, 9}}.
\medskip

To find this set, we first consider the set of \textit{all} update log records:
this is \ms{{3, 4, 5, 6, 9}}. We then remove the record with \ms{LSN = 6}, since
the affected page exists in the DPT with \ms{recLSN = 5}, which is lower than
\ms{6}.

\begin{itemize}
  \item \textit{5. The set of log records undone during the undo phase.}
\end{itemize}

The complete set of log records undone during the undo phase is \ms{{9, 4, 3}}
(in that order).
\medskip

Here's why: Starting with the \ms{lastLSN}s for \ms{T1} (given by the
transaction table), the undo phase traces log records backwards in time via the
\ms{prevLSN}s and \ms{undoNextLSN}s of update records and CLRs.
\streg

\begin{itemize}
  \item \textit{6. The contents of the log after the recovery procedure
    completes.}
\end{itemize}

\begin{minted}{text}
LSN | prevLSN  | Xact_ID | Type       | pageId   | undoNextLSN
----+----------+---------+------------+----------+------------
1   | -        | -       | begin CKPT | -        | -
2   | -        | -       | end CKPT   | -        | -
3   | NULL     | T1      | update     | P2       | -
4   | 3        | T1      | update     | P1       | -
5   | NULL     | T2      | update     | P5       | -
6   | NULL     | T3      | update     | P5       | -
7   | 6        | T3      | commit     | -        | -
8   | 5        | T2      | commit     | -        | -
9   | 4        | T1      | update     | P5       | -
10  | 7        | T3      | end        | -        | -
----------------------CRASH AND RESTART-----------------------
11  | 8        | T2      | end        | -        | -
12  | 9        | T1      | CLR        | P5       | 4
13  | 12       | T1      | CLR        | P1       | 3
14  | 13       | T1      | CLR        | P2       | NULL
15  | 14       | T1      | end        | -        | -
\end{minted}

Notice the \ms{end} record for \ms{T2}, which was written during the analysis
phase.

\tykstreg
