\iffalse{
  >> LEARNING GOALS

  % (L4A) Explain the theoretical basics behind CNNs
  % (L4B) Implement and test a specific CNN for image segmentation
  % (L4C) Define what cross-validation is
  % (L4D) Assess classifier performance with different metrics
  % (L4E) Design your own algorithm (e.g. CNN) and evaluate its performance on the skin lesion classification data.

  >> DISPOSITION

  >> 1: Convolutional Neural Networks in MIA (L4A)
  * what is a neural network? an ML model comprised of a network of connected
  \emph{neurons}, organized in so-called \emph{layers}, each with its own
  associated weights and biases -- much like a set of equations. Each neuron
  outputs a value that is fed to an activation function which indicates to the
  next layer whether or not the output of this particular neuron is important.

  section 1.1 convolutional NN
  * what is a convolutional neural network? One big problem arises when we
  attempt to use ANN's with image data: Input image data can vary greatly, and
  ideally we want to build a model that is robust regardless of the placement
  and orientation of objects in the image. This is called translation
  invariance. A CNN comes close by feeding the input image through consecutive
  convolution layers, each layer producing multiple feature maps, each map
  capturing different features of the image -- for example, in one layer we
  might have a set of neurons in a given layer which solely detect diagonal
  rounded edges, and another set which boosts and captures the contrast between
  foreground and background objects -- and then feeding the result of this
  through downsampling (or pooling) layers, which extract defining features and
  compresses the input into an exponentially smaller map of information.

  This downsampling greatly reduces computational overhead as opposed to a
  regular ANN, however, a natural consequence, of course, is that we must trade
  off translation invariance for -equivariance (ie. the output \emph{is}
  affected by translation of the input, but the translation is similar and thus
  predictable).


  section 2: U-Net and implementation (L4B + L4C?)

  describe the U-Net model for image segmentation.
  contracting and expanding path.
  idea is to extract features, then upsample the extracted features to something
  resembling the input image, but just the important features.
  example of such a network (graph, or just textual description????)

  show example segmentation (and mention why it fails for that one stupid mask)

  subsection 2.1 : training and cross-validation (L4C)

  subsection 2.2: evaluation (L4D)
  loss/accuracy graph. mention that loss seems to still be decreasing, which
  might indicate a necessity for further fitting, however, validation loss seems
  to have plateaued.
  ROC graph and AUC. explain why AUC = 0.97 is pretty good.


  section 3: skin lesion image binary classification (L4E)

  very short description of dataset: benign/malignant. We want a binary
  classification of the data cases.

  subsection 3.1:
  describe the model: same as prior U-Net model, but with a dense layer substituted
  for the entire expanding path.

  subsection 3.2: Regularization using data augmentation
  The supplied training data is heavily imbalanced with 80.8\% benign cases.
  With the data as is, the model is very likely to simply always predict
  ''benign``, for a training accuracy of $\sim 80.8\%$. In cases such as this
  one, data augmentation is a suitable regularization method, can be used to
  straighten out the imbalance. Hence, I augment the dataset with 522 malignant
  cases for a total of 695, to better match the 727 benign cases, using random
  rotation, translation, noise/blur, and brightness/contrast.

  subsection 3.x: classifier assessment

  subsubsection: assessing training process

  show loss/accuracy of training process. explain that this can be used to
  assess whether model tweaking or more fitting is needed (decreasing loss can
  indicate need for more fitting; small loss but unexpected accuracy can
  indicate bad local minimum, which can further indicate overly complex model or
  not enough training data).

  subsubsection: assessing results

  After training: run on test data.
  receiver operating characteristic curve
}\fi


\documentclass{llncs}

\input{commands_and_packages.tex}

\title{MIA Assignment 4}
\author{Anders Holst (wlc376)}
\institute{}

\begin{document}

\pagenumbering{arabic}
\maketitle

\input{1.tex}
\input{2.tex}
\input{3.tex}

\newpage

% \begin{thebibliography}{8}

  % \bibitem{ronneberger}
  %   Ronneberger, Olaf; Fischer, Philipp; Brox, Thomas. \textit{U-Net:
  %   Convolutional Networks for Biomedical Image Segmentation}. arXiv:1505.04597.
  %   \url{https://arxiv.org/abs/1505.04597}.
  % \bibitem{wu}
  %   Wu, Zufeng; Lan, Tian; Wang, Jiang; Ding, Yi; Qin, Zhiguang. \textit{Medical
  %   Image Registration Using B-Spline Transform}. School of Information and
  %   Software Engineering, University of Electronic Science and Technology of
  %   China.
  %
  %   \url{https://ijssst.info/Vol-17/No-48/paper1.pdf}.

% \end{thebibliography}

\end{document}
