% \documentclass[a4paper,12pt]{article}
%
% \input{commands_and_packages.tex}
%
% \newcommand{\concat}{\ensuremath{+\!\!\!\!+\,}}
% \newcommand{\rpm}{\raisebox{.2ex}{$\scriptstyle\pm$}}
% \newcommand{\resub}[1]{\textcolor{red}{#1}}
% \newcommand{\blue}[1]{\textcolor{blue}{#1}}
% \newcommand{\ltt}{\sqsubseteq}
% \newcommand{\twodots}{\mathinner {\ldotp \ldotp}}
%
% \title{\large MIA -- Hand-in 3}
% \author{{\footnotesize Anders Lietzen Holst (wlc376)}\\ {\footnotesize \today}}
%
% \date{}

% % (L3C) Describe regularization.

% % (L3E) Implement a non-rigid registration in python, including a
% %         transformation model and a similarity measure.
%
% % (L3B) Describe the role of similarity measure.
% % (L3A) Describe what the registration problem is.
% % (L3D) Knowledge of the different types of transformations.


\iffalse{
  >> disposition

  section 1:
  * describe registration -- geometrically transform an image into (something
      geometrically resembling) another. useful tool to be able to compare
      different images of the same thing (or the same thing in different
      bodies). For example, if two images are created using the same modality
      (x-ray, MRI, etc) and are aligned wrt. the subject, determining the
      correlation may not be a problem at all, but sometimes we may eg. want to
      compare an x-ray to an MRI, or to compare an image of one body to the
      image of another, perhaps taken with no knowledge of the first image.

  * when is it used and why
  * supervised algorithm; needs training data


  section 2:
  steps of the algorithm -- in particular, describe similarity:

    pixelwise similarities (eg. MSE) and (normalized) cross correlation match
    pixel intensities, but we often want to match \textit{patterns}: hence
    mutual image information, based on joint entropy.

    The mutual information $MI(I, J)$ measures the amount of information one can
    obtain about an image $I$ from the image $J$, and vice versa, and is given
    by the double sum over image pixels:


\begin{align*}
  MI(\mathcal I, \mathcal J) &= \sum_{i \in \mathcal I} \sum_{j \in \mathcal J}
  p(i, j) \log{3}\TODO{this}
\end{align*}

  and an image registration is finished when the mutual information is
  \textit{maximized}.


  section 3:
  * different types of transformations: rigid, non-rigid

  * regularization?

}\fi



\documentclass{llncs}

\input{commands_and_packages.tex}

\title{Mia Assignment 3}
\author{Anders Holst (wlc376)}
\institute{}

\begin{document}

\pagenumbering{arabic}
\maketitle

\input{1.tex}
\input{2.tex}


\begin{thebibliography}{8}
  \bibitem{wu}
    Wu, Zufeng; Lan, Tian; Wang, Jiang; Ding, Yi; Qin, Zhiguang. \textit{Medical
    Image Registration Using B-Spline Transform}. School of Information and
    Software Engineering, University of Electronic Science and Technology of
    China.

    \url{https://ijssst.info/Vol-17/No-48/paper1.pdf}.

\end{thebibliography}

\end{document}
